---
title: "HW3"
author: "Miles Taylor"
date: "6/10/2022"
output: pdf_document
---

```{r setup, echo= FALSE, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(kableExtra)
library(reticulate)
knitr::opts_chunk$set(echo = TRUE)

df <- read.csv('veneer.csv')
```

**Do a boxplot(brand ~ wear)**

$~$

```{r}

ggplot(data  = df) + 
  geom_boxplot(aes(x=BRAND, y=WEAR, fill = BRAND)) + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  xlab('Brand') +
  ylab('Wear') +
  ggtitle('Wear by Brand') + 
  theme_bw()
                 

                   
```

$~$


**Use anova to test that the mean of the 4 brands are the same.**

$~$

```{r}
anova <- aov(df$WEAR ~ factor(df$BRAND))

sm <- summary(anova)
```

```{python}
import pandas as pd

output_df = r.sm[0]

```

The output from the ANOVA model is specified below: 

```{r}
kbl(py$output_df)
```

$~$

Tukey's HSD is graphed below to confirm the result

```{r}
plot(TukeyHSD(aov(df$WEAR ~ factor(df$BRAND))))



```

$~$

**Conclusion**

The null hypothesis H0 that the means of the BRANDS are the same should
be rejected. The ANOVA test shows a significant difference for the means at
the 95% confidence level (even at the 99% confidence interval) through
analysis of the P-value 0.001683 < 0.05.   

This test is only guarantees that at least mean from one group, say ACME,
is significantly different than the rest of the data set, but says 
nothing about which group. Tukey's HSD is performed adhoc to 
confirm the result visually and show the spread of the means of the groups. ACME, TUFFY,
and XTRA have insignificant differences between the means of their groups(you can
inspect this visually by looking at the IQR for the box and whiskey plot).


